Computing exact restricted isometry constants via\\ mixed integer semidefinite programming

### 2016-03-09 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
reworked all parts of the paper except for the numerical results

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{mathtools}

@@

One important condition is the \emph{restricted isometry property} (RIP),
which was first introduced by Cand\`es and Tao~\cite{CT05}. For each $k \in
\{1, \dots, n\}$, the RIP involves a \emph{restricted isometry constant}
$\delta_k$ bounding the $k$-sparse eigenvalues of $A$. If $\delta_k$ is sufficiently small
and there exists $\tilde{x} \in \R^n$ with $A \tilde{x} = b$ and
$\Norm{\tilde{x}}{0} \leq k$, then problems~\eqref{l0min} and~\eqref{l1min}
are equivalent; see Section~\ref{sec:RIP} for more details. There are
several types of matrices~$A$ for which the RIP holds, for instance random
matrices, see, e.g., Baraniuk et al.~\cite{BDDW08}. Although the RIP does
not provide necessary conditions, it is one of the major tools for
establishing for $\ell_0$-$\ell_1$-equivalence.

@@

The goal of this paper is to compute optimal restricted isometry constants
via a mixed integer semidefinite program (SDP). To this end, we will
introduce an SDP with binary variables and a rank-1 constraint. In order to
utilize this SDP, we will show that the rank constraint can be relaxed
without changing the optimal value. Moreover, we discuss a solution
methodology to solve the resulting mixed integer SDP. It will turn out that
the resulting problems are quite hard to solve, even for matrices of
smaller size. However, the results allow an evaluation of the SDP
relaxations of d'Aspremont et al. Depending on the input matrices, these
bounds can be quite tight or weak. This complements the theoretical
approximation guarantees developed by d'Aspremont et
al. in~\cite{AspBG14}.

In the following, we will start by introducing the restricted isometry
property and corresponding results about the equivalence of~\eqref{l0min}
and~\eqref{l1min} in Section~\ref{sec:RIP}. In Section~\ref{sec:SDPrelax},
we will then review the SDP-relaxations of~\cite{Asp07}
and~\cite{Asp08}. Afterwards we will construct a mixed integer semidefinite
programme and show that it produces exact restricted isometry constants in
Section~\ref{sec:MISDP}. Finally, in Section~\ref{sec:numerical_results},
we present numerical results and comparisons with the SDP relaxations in
\cite{Asp07} and \cite{Asp08}.

@@

\begin{definition}\label{ricdef}
For a matrix $A \in \R^{m \times n}$ and order~$k$, the smallest constant in 
(\ref{rip}), meaning
\begin{equation}\label{ric}
 \delta_k \define \argmin_{\delta \geq 0} \{(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\; \text{ for all }x \in \Sigma_k\},
\end{equation}
is called \emph{restricted isometry constant} (RIC).
\end{definition}

@@

\begin{definition}\label{asymRIPdef}
 The \emph{lower} and \emph{upper RIC} are defined as
 \begin{align}
   \alpha_k & \define \argmax_{\alpha \geq 0}\; \{ \alpha^2 \Norm{x}{2}^2
   \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\}\text{ and }\label{lowerRIP}\\
   \beta_k & \define \argmin_{\beta \geq 1}\; \{ \Norm{Ax}{2}^2 \leq \beta^2
   \Norm{x}{2}^2,\; \text{ for all }x \in \Sigma_k\},\label{upperRIP}
 \end{align}
 respectively. The \emph{restricted isometry ratio} (RIR) $\gamma_k$ is
 defined as $\gamma_k \define \beta_k^2 / \alpha_k^2$.
\end{definition}

\comment{Der ganze Absatz ist \"uberarbeitet, z.B.\ die RIR korrigiert
  (statt $(1 - \delta_k)/(1 + \delta_k)$). Korrekt? Korrekt}
%
Definition~\ref{asymRIPdef} is a generalization of Definition~\ref{ripdef}
in the following sense: if the matrix~$A$ satisfies the RIP of order $k$
with RIC $\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
\delta_k)$. Conversely, if $A$ has lower/upper RICs $\alpha_k$ and
$\beta_k$, respectively, then this implies a restricted isometry constant
of $\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ in~\eqref{rip}.
\comment{$\max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ erf\"ullt RIP nicht nur,
sondern ist auch das kleinste $\delta$, das es tut, weshalb die asym. wirklich
st\"arker ist.} So $A$ has a small
RIC if and only if it has a large lower and a small upper RIC. But since
in~\eqref{rip} usually only one of the two inequalities will be sharp, the
asymmetric version yields more information about~$A$.

@@

Again this result can be generalized to the denoising case. One advantage
of Defintion~\ref{asymRIPdef} and Theorem~\ref{asymRIPtheorem} is that they
are invariant under scaling of $Ax = b$. The condition of
Theorem~\ref{RIPsqrt2} holding for~$A x = b$ might not hold for the scaled
$\lambda Ax = \lambda b$. However, in Definition \ref{asymRIPdef}, we can
use $\tilde{\alpha}_k = \lambda \alpha_k$, $\tilde{\beta}_k = \lambda
\beta_k$, and therefore $\tilde{\gamma}_k = \lambda^2 \alpha_k^2 /
\lambda^2 \beta_k^2 = \gamma_k$.

@@

In this section we will discuss two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in (\ref{rip}), as they have some ideas in common with the mixed integer semidefinite program we want to 
propose and will also be checked against in the section on numerical results. The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k^2$ in (\ref{lowerRIP}) can be computed
with the non-convex quadratic optimization problem
\begin{align}\label{QP}
 \min \quad & \Norm{Ax}{2}^2 \nonumber \\
 \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
 & \Norm{x}{0} \leq k. \nonumber
\end{align}
The corresponding maximization problem can be solved to compute $\beta_k^2$. In \cite{Asp07} the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95} 
for the max-cut problem. The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where 
$X \succeq 0$ means that $X$ is symmetric (shortly written \mbox{$X \in S_n$}) and positive semidefinite. The objective and the normalization-constraint can then be expressed in $X$ as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
\Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$. In the
next step the inequality $\Norm{x}{0} \leq k$ is then substituted by the coarser constraint $\ones\T
\abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}} \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank 1 case, where $\vect(X)$ is a vector
in $\R^{n^2}$ consisting of all entries of $X$. 
Relaxing the non-convex Rank-constraint finally leads to the following relaxation:
\begin{align}\label{Asp07}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
 & \ones\T \abs{X} \ones \leq k \tag{A1} \\
 & X \succeq 0. \nonumber
\end{align}

Another SDP relaxation was proposed in \cite{Asp08}. There the problem

\begin{equation}\label{phi}
 \phi(\rho) = \max_{\Norm{x}{2} \leq 1} x\T \Sigma x - \rho \Norm{x}{0},
\end{equation}
originally stemming from sparse principal component analysis, is tackled for a symmetric positive semidefinite matrix $\Sigma$. For
$\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrange-relaxation as
\begin{equation*}
\beta_k^2 = \max_{\|x\|_2 \leq 1, \|x\|_0 \leq k} x^TA^TAx \leq \inf_{\rho \geq 0} \max_{\|x\|_2\leq 1} x^TA^TAx - \rho (\|x\|_0-k) = \inf_{\rho \geq 0} \phi(\rho) + \rho k.
\end{equation*}

As one way to solve (\ref{phi}) again the semidefinite lifting technique is used in \cite{Asp08} to compute an upper bound on $\phi$ via
\begin{align}\label{Asp08}
 \max \quad & \sum_{i=1}^n\Tr(P_iB_i) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
 & X \succeq 0 \tag{A2-Primal} \\
 & X \succeq P_i \succeq 0 \quad \text{for } i = 1, ..., n, \nonumber
\end{align}
for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. The corresponding dual problem can then be written as

\begin{align}\label{Asp08Dual}
  \min \quad & \lambda_{\max}\left(\sum_{i=1}^n Y_i \right) \nonumber \\
  \text{s.t.} \quad & Y_i \succeq B_i \quad \forall i \leq n \tag{A2-Dual} \\
  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, ..., n, \nonumber
\end{align}
which only has half as many variables as (\ref{Asp08}) and one less linear and semidefinite constraint. Then (\ref{Asp08Dual}) can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
optimal solution for (\ref{phi}) will always be $x=0$.

As mentioned in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
\begin{equation*}
 \alpha_k^2 \geq \sup_{\rho \geq 0} \psi(\rho) - \rho k,
\end{equation*}
with 
\begin{equation*}
 \psi(\rho) = \min_{\Norm{x}{2} \leq 1} x\T \Sigma x + \rho \Norm{x}{0}.
\end{equation*}
In this case problem (\ref{Asp08}) or equivalently (\ref{Asp08Dual}) has to be solved for $\Sigma = M I - A\T A$, where $M$ has to be big enough to make $\Sigma$ positive semidefinite again. If $\bar{\Psi}$ is the corresponding optimal solution, this implies
a lower bound on $\alpha_k$ as
\begin{equation*}
 \alpha_k^2 \geq \sup_{\rho \geq 0} (M - \bar{\Psi}) - \rho k.
\end{equation*}

@@

In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{lowerRIP}) and (\ref{upperRIP}) using mixed integer semidefinite programming. We will again start with the non-convex quadratic problem
\begin{align}\label{QP2}
 \min \quad & \Norm{Ax}{2}^2 \nonumber \\
 \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
 & \Norm{x}{0} \leq k. \nonumber
\end{align}

For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$. We can thus rewrite (\ref{QP}) equivalently to
% In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}) using mixed integer semidefinite programming. A non-convex quadratic optimization problem to compute $\alpha_k$ can be formulated as
% 
% \begin{align}\label{QP}
%  \min \quad & \Norm{Ax}{2}^2 \nonumber \\
%  \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
%  & \Norm{x}{0} \leq k. \nonumber
% \end{align}
% 
% By switching $\min$ for $\max$ in (\ref{QP}) we also get a corresponding optimization problem to compute $\beta_k$ and therefore also $\gamma_k$ and $\delta_k$.
% The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$. For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
% use the technique of semidefinite lifting first introduced in \cite{GW95} for the max-cut problem. The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the 
% equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where $X
% \succeq 0$ means that $X$ is symmetric and positive semidefinite. Using
% that $\Norm{Ax}{2}^2 = x\T A\T Ax = \Tr(A\T A X)$ and 
% $\Tr(X) = \sum_{i=1}^n x_i^2$, we can thus rewrite (\ref{QP}) equivalently to
\begin{align}\label{Rk1MISDP}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, ..., n \nonumber \\
 & \sum_{i=1}^n z_i \leq k \tag{M1} \\
 & \Rk(X) = 1 \nonumber \\
 & X \succeq 0  \nonumber \\
 & z \in \{0,1\}^n. \nonumber
\end{align}
\comment{test whether to use only diagonal inequalities $-z_j \leq X_{jj} \leq z_j$, or only those for $j$ (which are redundant in the SDP because of diagonal dominance) or all for $i$ and $j$ (which are even more redundant 
because of symmetrie)}

The next step would then be to relax the non-convex rank constraint again to arrive at the following relaxation which computes a lower bound on the optimal constant $\alpha_k$ in (\ref{lowerRIP}):

\begin{align}\label{MISDP}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, ..., n \nonumber \\
 & \sum_{i=1}^n z_i \leq k \tag{MISDP} \\
 & X \succeq 0  \nonumber \\
 & z \in \{0,1\}^n \nonumber
\end{align}

@@

Now let $\bar{X}$ be the $\ell \times \ell$ matrix with all rows and columns outside of $T$ removed. Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), 
as we only removed zero rows and columns of $\hat{X}$ and $(A\T A)\hat{X}$, which do not influence the trace or the positive semidefiniteness.

On the other hand we can lift any optimal solution $\check{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $z^{\prime}_i$ to one if and only if $i \in T$
and $X^{\prime}_{i_{j_1} i_{j_2}} = \check{X}_{j_1j_2}$ if and only if $i_{j_1}, i_{j_2} \in T$
and $X^{\prime}_{ij}=0$ otherwise. This solution will then be feasible for (\ref{MISDP}), as $\Tr(\check{X}) = 1$ and $\check{X}\in S_\ell$ imply $0 \leq \check{X}_{jj} = X^{\prime}_{i_j i_j} \leq 1$ and therefore 
$X^{\prime}_{ij} \leq z^{\prime}_j$ by diagonal dominance of semidefinite matrices.

So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 
$\Rk(X^{\prime}) = \Rk(\check{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$, such a point exists as a consequence of corollary 18.5.1 in
\cite{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using theorem \ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
\begin{equation*}
 \frac{1}{2}\Rk(\breve{X})\cdot(\Rk(\breve{X})+1) \leq \tilde{m} + \textbf{dim}(F) = 1 + 0 = 1.
\end{equation*}
As $\breve{X}=0$ is infeasible for $\Tr(X)=1$, we must have $\Rk(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
\end{proof}

As Theorem \ref{Rk1thm} tells us that (\ref{MISDP}) always has a solution of rank one and the rank-constraint was the only constraint we relaxed to get (\ref{MISDP}) from (\ref{QP}), this means that their optimal objective values 
have to agree and we can compute the optimal constant $\alpha_k$ for the RIP by solving (\ref{MISDP}). As the same argumentation also holds for the maximization problem for $\beta_k$, this allows us to compute the restricted isometry
constant and the restricted isometry ratio by mixed integer semidefinite programming.

@@

Theorem \ref{Rk1thm} and Remark \ref{Rk1rmk} do not guarantee that any found solution will have rank one, they only guarantee that there exists a rank one solution. But this isn't a problem for applications in the context of the RIP, 
as we are here only interested in the optimal objective value, as this gives us the constant $\alpha_k$ or $\beta_k$ for the RIP. For other applications like the SPCA such an extreme solution can also be generated with an algorithm 
proposed by Pataki in \cite{coneLP}.

\section{Solving MISDPs}
\label{sec:solving_misdps}

>>>>

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{url}

@@

One important condition is the \emph{restricted isometry property} (RIP),
which was first introduced by Cand\`es and Tao~\cite{CT05}. For each $k \in
\{1, \dots, n\}$, the RIP involves a \emph{restricted isometry constant}
$\delta_k$ bounding the $k$-sparse eigenvalues of $A$. If $\delta_k$ is sufficiently small
and there exists $\tilde{x} \in \R^n$ with $A \tilde{x} = b$ and
$\Norm{\tilde{x}}{0} \leq k$, then problems~\eqref{l0min} and~\eqref{l1min}
are equivalent; see Section~\ref{sec:RIP} for more details. There are
several types of matrices~$A$ for which the RIP holds, for instance random
matrices, see, e.g., Baraniuk et al.~\cite{BDDW08}. Although the RIP does
not provide necessary conditions, it is one of the major tools for
establishing the $\ell_0$-$\ell_1$-equivalence.

@@

The goal of this paper is to compute optimal restricted isometry constants
via a mixed integer semidefinite program (MISDP). To this end, we will
introduce an MISDP with a rank-1 constraint. In order to
utilize this MISDP, we will show that the rank constraint can be relaxed
without changing the optimal value. Moreover, we discuss a solution
methodology to solve the resulting MISDP. It will turn out that
the resulting problems are quite hard to solve, even for matrices of
smaller size. However, the results allow an evaluation of the SDP
relaxations of d'Aspremont et al. Depending on the input matrices, these
bounds can be quite tight or weak. This complements the theoretical
approximation guarantees developed by d'Aspremont et
al. in~\cite{AspBG14}.

In the following, we will start by introducing the restricted isometry
property and corresponding results about the equivalence of~\eqref{l0min}
and~\eqref{l1min} in Section~\ref{sec:RIP}. In Section~\ref{sec:SDPrelax},
we will then review the SDP-relaxations of~\cite{Asp07}
and~\cite{Asp08}. Afterwards we will construct a mixed integer semidefinite
programm and show that it produces exact restricted isometry constants in
Section~\ref{sec:MISDP}. Finally, in Section~\ref{sec:numerical_results},
we present numerical results and comparisons with the SDP relaxations in
\cite{Asp07} and \cite{Asp08}.

@@

\begin{definition}\label{ricdef}
For a matrix $A \in \R^{m \times n}$ and order~$k$, the smallest constant in 
(\ref{rip}), meaning
\begin{equation}\label{ric}
 \begin{aligned}
  \delta_k \define \argmin_{\delta \geq 0} \{&(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\\
  & \text{ for all }x \in \Sigma_k\},
 \end{aligned}
\end{equation}
is called \emph{restricted isometry constant} (RIC).
\end{definition}

@@

\begin{definition}\label{asymRIPdef}
 The \emph{lower} and \emph{upper RIC} are defined as
 \begin{align}
   \alpha_k & \define \argmax_{\alpha \geq 0}\; \{ \alpha^2 \Norm{x}{2}^2
   \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\}, \label{lowerRIP}\\
   \beta_k & \define \argmin_{\beta \geq 1}\; \{ \Norm{Ax}{2}^2 \leq \beta^2
   \Norm{x}{2}^2,\; \text{ for all }x \in \Sigma_k\},\label{upperRIP}
 \end{align}
 respectively. The \emph{restricted isometry ratio} (RIR) $\gamma_k$ is
 defined as $\gamma_k \define \beta_k^2 / \alpha_k^2$.
\end{definition}

%
Definition~\ref{asymRIPdef} is a generalization of Definition~\ref{ripdef}
in the following sense: if the matrix~$A$ satisfies the RIP of order $k$
with RIC $\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
\delta_k)$. Conversely, if $A$ has lower/upper RICs $\alpha_k$ and
$\beta_k$, respectively, then this implies a restricted isometry constant
of $\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ in~\eqref{rip}.
So $A$ has a small RIC if and only if it has a large lower and a small upper 
RIC. But since in~\eqref{rip} usually only one of the two inequalities will 
be sharp, the asymmetric version yields more information about~$A$.

@@

Again this result can be generalized to the denoising case. One advantage
of Definition~\ref{asymRIPdef} and Theorem~\ref{asymRIPtheorem} is that they
are invariant under scaling of $Ax = b$. The condition of
Theorem~\ref{RIPsqrt2} holding for~$A x = b$ does not imply it for the scaled
$\lambda Ax = \lambda b$. However, in Definition \ref{asymRIPdef}, we can
use $\tilde{\alpha}_k = \lambda \alpha_k$, $\tilde{\beta}_k = \lambda
\beta_k$, and therefore $\tilde{\gamma}_k = \lambda^2 \alpha_k^2 /
\lambda^2 \beta_k^2 = \gamma_k$.

@@

In this section we will discuss two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in (\ref{rip}), as they have some ideas in common with the mixed integer semidefinite program we want to 
propose and will also be checked against in the section on numerical results. The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k^2$ in (\ref{lowerRIP}) can be computed
by the non-convex quadratic optimization problem
\begin{align}\label{QP}
 \min \quad & \Norm{Ax}{2}^2 \nonumber \\
 \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
 & \Norm{x}{0} \leq k. \nonumber
\end{align}
The corresponding maximization problem can be solved to compute $\beta_k^2$. In \cite{Asp07} the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95} 
for the max-cut problem. The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where 
$X \succeq 0$ means that $X$ is symmetric (shortly written \mbox{$X \in S_n$}) and positive semidefinite. The objective and the normalization-constraint can then be expressed in $X$ as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
\Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$. In the
next step the inequality $\Norm{x}{0} \leq k$ is then substituted by the coarser constraint $\ones\T
\abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}} \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank 1 case, where $\vect(X)$ is a vector
in $\R^{n^2}$ consisting of all entries of $X$. 
Relaxing the non-convex Rank-constraint finally leads to the following relaxation:
\begin{align}\label{Asp07}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
 & \ones\T \abs{X} \ones \leq k \tag{A1} \\
 & X \succeq 0. \nonumber
\end{align}

Another SDP relaxation was proposed in \cite{Asp08}. There the problem
\begin{equation}\label{phi}
 \phi(\rho) = \max_{\Norm{x}{2} \leq 1} x\T \Sigma x - \rho \Norm{x}{0},
\end{equation}
originally stemming from sparse principal component analysis, is tackled for a symmetric positive semidefinite matrix $\Sigma$. For
$\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrangian Relaxation as
\begin{equation*}
 \begin{aligned}
  \beta_k^2 & = \max_{\|x\|_2 \leq 1, \|x\|_0 \leq k} x^TA^TAx \\
	    & \leq \inf_{\rho \geq 0} \max_{\|x\|_2\leq 1} x^TA^TAx - \rho (\|x\|_0-k) \\
	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho k.
 \end{aligned}
\end{equation*}

As one way to solve (\ref{phi}) again the semidefinite lifting technique is used in \cite{Asp08} to compute an upper bound on $\phi$ via
\begin{align}\label{Asp08}
 \max \quad & \sum_{i=1}^n\Tr(P_iB_i) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
 & X \succeq 0 \tag{A2-Primal} \\
 & X \succeq P_i \succeq 0 \quad \text{for } i = 1, ..., n, \nonumber
\end{align}
for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. The corresponding dual problem can then be written as 
\begin{align}\label{Asp08Dual}
  \min \quad & \lambda_{\max}\left(\sum_{i=1}^n Y_i \right) \nonumber \\
  \text{s.t.} \quad & Y_i \succeq B_i \quad \forall i \leq n \tag{A2-Dual} \\
  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, ..., n, \nonumber
\end{align}
which only has half as many variables as (\ref{Asp08}) and one less linear and semidefinite constraint. Then (\ref{Asp08Dual}) can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
optimal solution for (\ref{phi}) will always be $x=0$.

As mentioned in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
\begin{equation*}
 \alpha_k^2 \geq \sup_{\rho \geq 0} \psi(\rho) - \rho k,
\end{equation*}
with 
\begin{equation*}
 \psi(\rho) = \min_{\Norm{x}{2} \leq 1} x\T \Sigma x + \rho \Norm{x}{0}.
\end{equation*}
In this case problem (\ref{Asp08}) or equivalently (\ref{Asp08Dual}) has to be solved for $\Sigma = M I - A\T A$, where $M$ has to be big enough to make $\Sigma$ positive semidefinite again. If $\bar{\Psi}$ is the corresponding optimal solution, this implies
a lower bound on $\alpha_k$ in
\begin{equation*}
 \alpha_k^2 \geq \sup_{\rho \geq 0} (M - \bar{\Psi}) - \rho k.
\end{equation*}

@@

In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{lowerRIP}) and (\ref{upperRIP}) using mixed integer semidefinite programming. We will again start with the non-convex quadratic problem
\begin{align}\label{QP2}
 \min \quad & \Norm{Ax}{2}^2 \nonumber \\
 \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
 & \Norm{x}{0} \leq k. \nonumber
\end{align}
For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$. We can thus rewrite (\ref{QP}) equivalently as
% In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}) using mixed integer semidefinite programming. A non-convex quadratic optimization problem to compute $\alpha_k$ can be formulated as
% 
% \begin{align}\label{QP}
%  \min \quad & \Norm{Ax}{2}^2 \nonumber \\
%  \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
%  & \Norm{x}{0} \leq k. \nonumber
% \end{align}
% 
% By switching $\min$ for $\max$ in (\ref{QP}) we also get a corresponding optimization problem to compute $\beta_k$ and therefore also $\gamma_k$ and $\delta_k$.
% The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$. For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
% use the technique of semidefinite lifting first introduced in \cite{GW95} for the max-cut problem. The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the 
% equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where $X
% \succeq 0$ means that $X$ is symmetric and positive semidefinite. Using
% that $\Norm{Ax}{2}^2 = x\T A\T Ax = \Tr(A\T A X)$ and 
% $\Tr(X) = \sum_{i=1}^n x_i^2$, we can thus rewrite (\ref{QP}) equivalently to
\begin{align}\label{Rk1MISDP}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, ..., n \nonumber \\
 & \sum_{i=1}^n z_i \leq k \tag{M1} \\
 & \Rk(X) = 1 \nonumber \\
 & X \succeq 0  \nonumber \\
 & z \in \{0,1\}^n. \nonumber
\end{align}
The next step would then be to relax the non-convex rank constraint again to arrive at the following relaxation which computes a lower bound on the optimal constant $\alpha_k$ in (\ref{lowerRIP}):
\begin{align}\label{MISDP}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, ..., n \nonumber \\
 & \sum_{i=1}^n z_i \leq k \tag{MISDP} \\
 & X \succeq 0  \nonumber \\
 & z \in \{0,1\}^n \nonumber
\end{align}

@@

Now let $\bar{X}$ be the $\ell \times \ell$ submatrix of $\hat{X}$ with all rows and columns outside of $T$ removed. Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), 
as we only removed zero rows and columns of $\hat{X}$ and $(A\T A)\hat{X}$, which do not influence the trace or the positive semidefiniteness.

On the other hand we can lift any optimal solution $\check{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $z^{\prime}_i$ to one if and only if $i \in T$
and $X^{\prime}_{i_{j_1} i_{j_2}} = \check{X}_{j_1j_2}$ if and only if $i_{j_1}, i_{j_2} \in T$
and $X^{\prime}_{ij}=0$ otherwise. This solution will then be feasible for (\ref{MISDP}), since $\Tr(\check{X}) = 1$ and $\check{X}\in S_\ell$ imply $0 \leq \check{X}_{jj} = X^{\prime}_{i_j i_j} \leq 1$ and therefore 
$X^{\prime}_{ij} \leq z^{\prime}_j$ by diagonal dominance of semidefinite matrices.

So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 
$\Rk(X^{\prime}) = \Rk(\check{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$, such a point exists
\cite[Corollary 18.5.1]{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using Theorem~\ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
\begin{equation*}
 \frac{1}{2}\Rk(\breve{X})\cdot(\Rk(\breve{X})+1) \leq \tilde{m} + \textbf{dim}(F) = 1 + 0 = 1.
\end{equation*}
Since $\breve{X}=0$ is infeasible for $\Tr(X)=1$, we must have $\Rk(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
\end{proof}

Since Theorem~\ref{Rk1thm} guarantees that (\ref{MISDP}) always has a solution of rank one and the rank-constraint was the only constraint we relaxed to get (\ref{MISDP}) from (\ref{QP}), this means that their optimal objective values 
have to agree and we can compute the optimal constant $\alpha_k$ for the RIP by solving (\ref{MISDP}). Since the same argumentation also holds for the maximization problem for $\beta_k$, this allows us to compute the restricted isometry
constant and the restricted isometry ratio by mixed integer semidefinite programming.

@@

To solve \eqref{MISDP} successfully using a branch-and-bound-approach, it is important to have strong semidefinite relaxations for fractional $z$ variables. 
Therefore we will further strengthen the constraint coupling the continuous and binary variables to force the nonzero entries of $z$ towards one in the
semidefinite relaxations.
